Comparing Early Evolution of Flames in X-ray Bursts in Two and Three Dimensions

### 2023-03-14 xrb3/paper.tex
address Kiran's comments

X-ray bursts (XRBs) result from thermonuclear burning of an accreted
H/He or He layer on a neutron star \citep{galloway:2017}.
Observations of brightness oscillations in the rise of the lightcurve
suggest that the burning begins localized and then spreads across the
neutron star \citep{bhattacharyya:2007}. {\color{red} The next sentence
is long and it was hard to read on my initial pass over it. Could consider
splitting the parenthetical note off as a separate sentence, and switching
to active voice (write ``and we need hydrodynamic simulations" and replace
``are needed" with ``in order").} 
This spreading is inherently
a multi-dimensional phenomenon, and hydrodynamic simulations that
resolve the reactive zone, use realistic reaction networks, and
capture the scales over which rotation is important are needed to
understand the flame propagation and nucleosynthesis (especially
rp-process in mixed H/He bursts, as discussed in \citealt{rpprocess}).
{\color{red} Could split paragraph and start next sentence with
"Here we show...".}
We show a first attempt at modeling the early evolution of a spreading
hotspot in three dimensions.  This builds on our earlier two-dimensional
work \citep{eiden:2020,harpole:2021} that developed our
simulation framework and explored how the acceleration of the burning
front depends on the initial model structure.  As with those
simulations, we use the freely-available \castro\ simulation code
\citep{castro,castro_joss}, with all of the code needed to run these
simulations in our public GitHub repositories.

@@

Moving to 3D is very expensive, so to further reduce the computational
expense, we use an aniostropic resolution, with 32 cm resolution laterally and 16
cm resolution vertically.  This is finer vertically than done in
\citet{harpole:2021}.  Because
the hotspot is placed in the center of the domain, the size of the
domain also becomes a constraint, and as a result, the evolution time
we can reach will become limited by the time it takes the burning to
approach the edge of the domain.  As a result, we focus here on the
early evolution. {\color{red} Very minor: could start previous sentence with
``consequently" or ``therefore" since ``as a result" appears in the previous sentence.}

@@

\begin{figure}[t]
\centering
%\epsscale{0.75}
\plotone{abar_top_stack}
%\epsscale{1.0}
\caption{\label{fig:vr_abar} Volume rendering of the three-dimensional XRB
simulation showing the mean molecular weight, $\bar{A}$.  The flame is viewed
from above. {\color{red} It might be helpful to have the time for each panel either
on the plots or in this caption.}
}
\end{figure}

@@

\begin{figure*}[t]
\centering
%\epsscale{0.75}
\plotone{flame_wave_1000Hz_25cm_smallplt174562_slice}
%\epsscale{1.0}
\caption{\label{fig:3dslice} A vertical slice through the three-dimensional
simulation {\color{red} at 40~ms} showing the entire domain from the side.  The
flame structure is clearly seen. {\color{red} The flame structure could probably
be made a little clearer if this only showed the lower $10^4~\mathrm{cm}$ of the
domain like the 2D plots. Could also try further reducing the peak $\bar{A}$ on the
colorbar (e.g. to 5.5 or 5.25) both here and in the 2D plots to bring out the flame
even more. I think the plots are still pretty clear with the current colorbar though.}}
\end{figure*}

>>>>

X-ray bursts (XRBs) result from thermonuclear burning of an accreted
H/He or He layer on a neutron star \citep{galloway:2017}.
Observations of brightness oscillations in the rise of the lightcurve
suggest that the burning begins localized and then spreads across the
neutron star \citep{bhattacharyya:2007}. 
This spreading is inherently
a multi-dimensional phenomenon, requiring hydrodynamic simulations that
resolve the reactive zone, realistic reaction networks, and
domain sizes that capture the scales over which rotation in order to
understand the flame propagation and nucleosynthesis.  Detailed
nuclear physics is especially important for understanding the 
rp-process in mixed H/He bursts, as discussed in \citealt{rpprocess}.

Here we show a first attempt at modeling the early evolution of a spreading
hotspot in three dimensions.  This builds on our earlier two-dimensional
work \citep{eiden:2020,harpole:2021} that developed our
simulation framework and explored how the acceleration of the burning
front depends on the initial model structure.  As with those
simulations, we use the freely-available \castro\ simulation code
\citep{castro,castro_joss}, with all of the code needed to run these
simulations in our public GitHub repositories.

@@

Moving to 3D is very expensive, so to further reduce the computational
expense, we use an aniostropic resolution, with 32 cm resolution
laterally and 16 cm resolution vertically.  This is finer vertically
than done in \citet{harpole:2021}.  Because the hotspot is placed in
the center of the domain, the size of the domain also becomes a
constraint, and as a result, the evolution time we can reach will
become limited by the time it takes the burning to approach the edge
of the domain.  Consequently, we focus here on the early evolution.

@@

\begin{figure}[t]
\centering
%\epsscale{0.75}
\plotone{abar_top_stack}
%\epsscale{1.0}
\caption{\label{fig:vr_abar} Volume rendering of the three-dimensional XRB
simulation showing the mean molecular weight, $\bar{A}$.  The flame is viewed
from above. The panels show the state at 10 ms, 20 ms, and 40 ms (from top to bottom).}
\end{figure}

@@

\begin{figure*}[t]
\centering
%\epsscale{0.75}
\plotone{flame_wave_1000Hz_25cm_smallplt174562_slice}
%\epsscale{1.0}
\caption{\label{fig:3dslice} A vertical slice through the three-dimensional
simulation {\color{red} at 40~ms} showing the entire domain from the side.  The
flame structure is clearly seen. }
\end{figure*}
